A quick user guide to the\\ \scine-\qcm\ software suite for \mol

### 2022-02-15 dmrg/doc/qcmaquis_manual.tex

\usepackage[nott]{kpfonts}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
% \usepackage[libertine]{newtxmath}
\usepackage{graphicx}% Include figure files
\usepackage{xcolor}
\usepackage{bm}% bold math
\usepackage[colorlinks=true]{hyperref} % web links
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\definecolor{citecolour}{rgb}{0,0.8,0.2}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour,citecolor=citecolour}
\usepackage{framed}
\usepackage{longtable}
\usepackage[bf,singlelinecheck=off]{caption}
\usepackage{amsmath}
\usepackage{listings}
%\lstset{breaklines=true}
\usepackage{multirow}
% \usepackage[top=2.25cm, bottom=2.25cm, left=2cm, right=2.25cm]{geometry}
\usepackage{booktabs}
\usepackage{chemformula}
\usepackage{placeins}
\usepackage[numbers,sort&compress]{natbib}
\setcounter{secnumdepth}{4}
% \renewcommand{\baselinestretch}{1.5}
\newcommand{\relv}{release version 3.0}
\newcommand{\mol}{\textsc{OpenMolcas}}
\newcommand{\qcm}{\textsc{QCMaquis}}
\newcommand{\scine}{SCINE}
\newcommand{\hostp}{\mol}
\newcommand{\molbuild}{\texttt{my-Molcas-build}}
\newcommand{\molsrc}{\texttt{my-Molcas-src}}
\newcommand{\qcmsrc}{\texttt{my-QCMaquis-src}}
\newcommand{\qcmbuild}{\texttt{my-QCMaquis-build}}
%\setlength{\parindent}{1.0cm}
%\setlength{\parskip}{1.00ex plus0.5ex minus0.5ex}
% \setlength{\abovecaptionskip}{2.00ex plus0.25ex minus0.25ex}
% \setlength{\belowcaptionskip}{\abovecaptionskip}
%\setlength{\textfloatsep}{3.0ex plus0.5ex minus0.5ex}
%\setlength{\intextsep}{3.0ex plus0.5ex minus0.5ex}

@@

% \usepackage{bibentry}
%% required for the bibentry package

@@

\lstset{numbers=left,numberstyle=\tiny,numberblanklines=true,basicstyle=\ttfamily\small,frame=single,showlines=true,flexiblecolumns=true,breaklines=true,postbreak={\llap{\textcolor{red}{$\lhook\joinrel\longrightarrow$\kern2.1em}}},escapechar=|}

\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\kwd}[1]{\texttt{#1}}

\title{A quick user guide to the\\ \scine-\qcm\ software suite for \mol}
\author{Leon Freitag, Sebastian Keller, Stefan Knecht, Yingjin Ma,\\ Christopher Stein and Markus Reiher}
\publishers{ETH Z\"urich, Laboratorium f\"ur Physikalische Chemie, Vladimir-Prelog-Weg 2,\\ CH-8093 Z\"urich}
\date{\today}

@@

\vspace{2cm}
\noindent We kindly request that, for reproducibility reasons, any use of the \qcm\ software suite for  density matrix renormalization group (DMRG) calculations in \mol\ that results in published material should cite the set-up steered by settings and warm-up procedures described in:
\begin{framed}
		\noindent {\textbf{Check for a preprint on \href{http://arxiv.org/}{arXiv.org} (to appear soon).}}\\
	\noindent Freitag~L.; Keller~S.; Knecht~S.; Lindh~R.; Ma~Y.; Stein~C. J.; Reiher~M. \emph{in preparation}.
\end{framed}

The DMRG calculations are then conducted with the software \qcm\ that requires a citation. It is described in the following paper:
\begin{framed}
%\noindent {\textbf{Check for the journal article asap on \href{http://arxiv.org/pdf/1510.02026.pdf}{arXiv.org}.}}\\
\noindent Keller,~S.; Dolfi,~M.; Troyer,~M.; Reiher,~M. \emph{J. Chem. Phys.}
  \textbf{2015}, \emph{143}, 244118, \href{https://doi.org/10.1063/1.4939000}{doi:10.1063/1.4939000}.
\end{framed}

\qcm\ builds upon the ALPS MPS project. The \href{http://alps.comp-phys.org/static/mps\_doc/index.html}{ALPS MPS codes} implement the DMRG algorithm for variational ground and low-lying excited state search as well as time evolution of arbitrary one- and two-dimensional models in a matrix-product-state representation. They have been developed at ETH Zurich by Michele Dolfi and Bela Bauer in the group of Matthias Troyer with contributions from Sebastian Keller and Alexandr Kosenkov and at the University of Geneva by Timoth{\'e}e Ewart and Adrian Kantian in the group of Thierry Giamarchi.

@@

ALPS is a general open-source framework for the description of strongly correlated many-particle systems.
\begin{framed}
B. Bauer, \textit{et al.} (ALPS Collaboration), The ALPS project release 2.0: open source software for strongly correlated systems, \textit{J. Stat. Mech.} \textbf{2011} P05001. \href{http://dx.doi.org/10.1088/1742-5468/2011/05/P05001}{http://dx.doi.org/10.1088/1742-5468/2011/05/P05001}.
\end{framed}

@@

\pagenumbering{arabic}
\section{Introduction to the \qcm\ Software Suite}

\subsection{Overview and Goals}
The computational cost of traditional full CI and full CI-based multiconfigurational calculations (such as CASSCF) scales exponentially with the number of active orbitals, thus calculations beyond a certain limit (e. g. 18 electrons in 18 orbitals for CASSCF) become prohibitively expensive. A possible approach to overcome the scaling limit is to employ methods such as density matrix renormalization group (DMRG) \cite{Schollwock_AnnPhys_densitymatrix_2011,Marti_ZPhysChem_Density_2010,Chan2011,Chan_JChemPhys_Matrix_2016} to solve the full CI problem approximately.

In DMRG, usually the wave function is represented in form of a matrix product state
(MPS). The \qcm\ software suite solves a full CI problem approximately by means of
an efficient optimization of an MPS wave function based on a second-generation DMRG algorithm \cite{Keller_JChemPhys_efficient_2015}.
The quantum-chemical operators are represented as matrix product operators (MPOs) which provides the necessary flexibility to accommodate Abelian and non-Abelian symmetries as well as the implementation of non-relativistic and relativistic quantum chemical Hamiltonians \cite{maquis-rel}, respectively,
in a unified framework. We have implemented the special unitary group of degree 2 (SU(2)) in the MPO representation of the non-relativistic Hamiltonian to ensure
spin conservation \cite{Keller_JChemPhys_Spinadapted_2016}.
The current implementation of \qcm\ allows for efficient full-CI-type calculations of active space sizes beyond capabilities (``CAS(18,18)") of standard CI approaches.

The \qcm\ software suite comes with a Fortran/Python-based interface \cite{interface} to the \mol\ \cite{Aquilante_JComputChem_Molcas_2016} framework, where we have implemented a state-specific and state-average DMRG self-consistent field (DMRG-SCF) algorithm, the possibility to include solvent effects in DMRG calculations, analytic gradients for state-specific calculations, state interaction in a nonorthogonal basis\cite{Knecht_JChemTheoryComput_nonorthogonal_2016} and n-electron valence state perturbation theory employing the DMRG reference wave function.\cite{Freitag_JChemTheoryComput_Multireference_2017} These developments allow for, among others, structure optimizations, spin-orbit coupling calculations, evaluation of transition densities between non-orthogonal states for electronic and magnetic properties, as well as to account for dynamic correlation by means of the NEVPT2 approach.

\begin{framed}
	\large{\textbf{Advice:} The break-even point wrt computational cost for a DMRG-CI/-SCF calculation compared to a ``traditional" CAS-CI/-SCF calculation is approximately reached for a CAS(14,14) space. For active spaces smaller than CAS(14,14) we recommend to choose a traditional \mol\ CAS approach.}
\end{framed}

\subsection{What features are included in this release?}\label{sec:features-v1.0}

@@

\subsection{Licenses}\label{sec:license}
The \qcm{}-\mol{} interface and the NEVPT2 program are licensed under GNU LGPL, version 2.1.
The \qcm{} software suite and the ALPS library are distributed under \href{http://alps.comp-phys.org/static/software/applications/LICENSE.txt}{ALPS Application License version 1.0} and the \href{http://alps.comp-phys.org/static/software/ALPS/LICENSE.txt}{ALPS Library License version 1.0}.
\clearpage
\newpage
\section{Software Requirements \& Download}\label{sec:install-req}

\subsection{Prerequisites}
In order to install either \qcm\ or the quantum chemistry package \mol\ with DMRG support through the \qcm\ software suite \cite{Keller_JChemPhys_efficient_2015,interface}, requires the following libraries/programs:

\begin{itemize}
 \item Git
%  \item Python version 2.\texttt{x} with \texttt{x}\ $\geq 7$, with SciPy and NumPy
 \item HDF5 (\url{http://www.hdfgroup.org/HDF5}) with Python2 bindings (\texttt{h5py})
 \item SZIP
 \item GNU Scientific Library GSL (\url{http://www.gnu.org/software/gsl/})
 \item CMake version $\geq$3.7 (\url{https://cmake.org/})
 \item Boost version $\geq$1.56 (\url{https://github.com/boostorg/boost})
 \item Python 3.x with \texttt{h5py} package installed\\
(optional, required only for MPSSI and distributed RDM calculations for NEVPT2 or the python tools)
\end{itemize}
\vspace{0.01cm}
\noindent
Please make sure that these libraries/programs are available and their location is visible in your \texttt{\$PATH}\ and
\texttt{\$LD\_LIBRARY\_PATH}, respectively.
Note that these libraries are \emph{NOT} part of the installation package of the \qcm\ software suite (see Section \ref{sec:installation}\
for further details).

@@

If you want to download and install a standalone version of \qcm{}, it may be obtained from \url{https://scine.ethz.ch/download/qcmaquis}. After downloading, please proceed to Section~\ref{sec:install-qcm-stand} for installation instructions.

\clearpage
\newpage
\section{Installation}\label{sec:installation}

% QCMaquis in Molcas
\subsection{Installation of the \mol-\qcm{} interface}\label{sec:install-qcm-mol}

% \noindent In the following steps \ref{subsubsec:download}-\ref{subsubsec:runtime}\ we describe how to successfully build and install the \mol\ program together with the \qcm\ software suite.
The installation of \qcm\ has been tested for different
operating systems and compiler/math libraries environments found in Section~\ref{subsec:supportedOS}.
While other combinations might work equally well they are \emph{not} officially supported.

\noindent The installation of the \qcm\ software suite within \mol\ will
comprise several libraries which are \emph{automatically downloaded and installed} during the \mol\ build
process provided that DMRG support within \mol\ is requested by the user.
The list of external libraries comprises:
\begin{itemize}
\item Optionally: the NEVPT2 program
\end{itemize}
All of the above libraries will be installed locally in the user-defined build folder \texttt{my-Molcas-build} of \mol.

\subsubsection{Download and build folder setup}
Download \mol{} from the official \mol{} GitLab repository:

@@

\noindent Create a build folder \texttt{\molbuild} -- note that this folder \emph{does not} necessarily have to be a subfolder
of \texttt{my-Molcas-src} -- and change to this new folder:
\begin{verbatim}
mkdir /path/to/my-Molcas-build && cd /path/to/my-Molcas-build
\end{verbatim}

\subsubsection{Configuration}\label{subsubsec:configure}

@@

\paragraph{Configuration with the GNU compiler suite}\label{sec:gnu-conf}$\;$\\

\noindent To configure \mol\ and \qcm\ with the GNU compiler suite type
\begin{verbatim}
FC=gfortran CC=gcc CXX=g++ cmake -DDMRG=ON -DLINALG=MKL \
  -DQCMaquis_NAME="Your Name" -DQCMaquis_EMAIL="your@email.com" \
  /path/to/my-Molcas-src
\end{verbatim}

\noindent where we assumed (\texttt{-DLINALG=MKL}) that the MKL libraries are available (recommended option).
If the MKL libraries are not available internal math libraries of \mol\ can be requested with \texttt{-DLINALG=Internal} (default option for \texttt{LINALG} in \mol) or \texttt{-DLINALG=Runtime} (to use a system-wide BLAS or LAPACK installation). Please also specify your name and e-mail address in the \texttt{-DQCMaquis\_NAME} and \texttt{-DQCMaquis\_EMAIL} CMake options, otherwise the download of \qcm{} will not work.

\begin{framed}
	\noindent\large{Note: We strongly recommend to configure \qcm\ (and \mol) with the Intel Math Kernel Library (MKL) to ensure the best numerical performance. See Section \ref{subsec:supportedOS}\ for further details.}
\end{framed}
% \paragraph{Configuration with the Intel compiler suite}\label{sec:intel-conf}$\;$\\
\begin{framed}
 \noindent Note: Intel compiler is currently not supported. Running a binary compiled with Intel compilers may result in segmentation faults. We are currently investigating the issue.
\end{framed}

\paragraph{Upgrading an existing \mol{} installation}\label{sec:build-upgrade}$\;$\\

If you're upgrading from an \mol{} installation compiled with an older release of \qcm{}, you need to add an additional command line flag when invoking \texttt{cmake}, namely
\begin{verbatim}
  -DQCMaquis_UPDATE=ON
\end{verbatim}

This flag activates the \qcm{} update for the current and all the following \texttt{cmake} runs. To switch off \qcm{} updates, \texttt{cmake} may be run with
\begin{verbatim}
  -DQCMaquis_UPDATE=OFF
\end{verbatim}

@@

\paragraph{Additional options}$\;$\\
Following additional options can be switched on by passing additional flags to \texttt{cmake}:
\begin{itemize}
 \item \texttt{-DNEVPT2=ON}: Downloads and compiles the NEVPT2 program and its interface with \mol\ and \qcm.
 \item \texttt{-DOPENMP=ON}: Enable OpenMP parallelization for MOLCAS and NEVPT2 (see Section \ref{sec:parallel}). Note that \qcm{} is built with OpenMP support \emph{regardless} of this flag.
\end{itemize}

Other options to enable/disable various options in \mol{} may be passed to \texttt{cmake} at this step. Please consult the \mol{} manual for details.
% \paragraph{Additional \mol{} options}$\;$\\
% By running
% \begin{verbatim}
%  ccmake /path/to/my-Molcas-src
% \end{verbatim}
% in the build directory you may obtain a text user interface where several options for building \mol{}, along with the ones mentioned above, may be conveniently set. Use \texttt{ccmake} to set other various options for \mol{}: for details please consult the \mol{} manual.
\begin{framed}
 \noindent Note: it is currently discouraged to install the \qcm{}-\mol{} interface together with the \textsc{Block} or \textsc{CheMPS2} interfaces available in \mol{}. This setup has not been tested and may lead to an unusable installation. A fix allowing different interfaces to be installed at the same time is currently under development.
\end{framed}

@@

\noindent Installing an MPI-parallel version of \mol\ with DMRG support is possible although
\qcm\ itself is by default shared-memory OMP but \emph{not} yet MPI-parallelized.
To configure \mol\ for an MPI-installation type
\begin{verbatim}
FC=mpif90 CC=mpicc CXX=mpiCXX cmake -DDMRG=ON -DLINALG=MKL \
  -DQCMaquis_NAME="Your Name" -DQCMaquis_EMAIL="your@email.com" \
  /path/to/my-Molcas-src
\end{verbatim}

@@

A shared-memory OMP parallelized version of \mol\ can be activated with the option
\texttt{-DOPENMP=ON}\ passed to \texttt{cmake} during the configuration step, for example
\begin{verbatim}
FC=... CC=... CXX=... cmake -DDMRG=ON -DLINALG=MKL \
  -DQCMaquis_NAME="Your Name" -DQCMaquis_EMAIL="your@email.com" \
  /path/to/my-Molcas-src
\end{verbatim}

@@

After a successful configuration, type
\begin{verbatim}
make
\end{verbatim}
\noindent or
\begin{verbatim}
make -j8
\end{verbatim}
to compile \mol\ (in parallel on 8 cores) and install all its components in the build folder \texttt{\molbuild}.
In the same folder \qcm\ as well as the required \textsc{Boost} and \textsc{Alps}
libraries will be downloaded and installed, respectively.
The installation process thus requires a working internet connection.

@@

After having successfully passed the \qcm\ installation step as indicated by the CMake message
\begin{verbatim}
[xxx%] Installation of QCMaquis, ALPS and Boost was successful!
\end{verbatim}
set up your \mol{} environment by setting \texttt{MOLCAS} environment variable
\begin{verbatim}
export MOLCAS=/path/to/my-MOLCAS-build
\end{verbatim}
and then set up the \qcm{} runtime environment by running the following shell script
\begin{verbatim}
source $MOLCAS/qcmaquis/bin/qcmaquis.sh
\end{verbatim}
You might want to include the above lines in your local \texttt{\$HOME/.bashrc} to set up your environment automatically every time your session is started.

@@

\qcm\ in \hostp\ is tested and officially supported for Linux (x86\_64)
and MacOS X (10.12.x and 10.13.x ) operating systems. We support GCC compilers starting from 5.x up to 7.x along with the MKL (Linux, x86\_64) and MKL/Accelerate (Mac OS X) math libraries, respectively, to successfully build the \qcm\ software suite within \hostp.
Note that other combinations might work equally well but they are not officially supported at present.

@@

To verify that your installation of \mol\ and \qcm\ was successful, run the following command
\begin{verbatim}
pymolcas verify qcmaquis
\end{verbatim}
\noindent The message
\begin{verbatim}
Verification has been completed
\end{verbatim}
\noindent indicates that all tests passed correctly and your installation is good to go for production work.
\vspace{2ex}

@@

\begin{itemize}
  \item For issues regarding usage of \qcm\ in \mol{}, open an issue in the \mol{} bug tracker at
    \url{https://gitlab.com/Molcas/OpenMolcas/issues}
  \item For issues regarding the usage of \qcm{} standalone version, please send an e-mail to \href{mailto:dmrg@phys.chem.ethz.ch}{dmrg@phys.chem.ethz.ch}.
\end{itemize}

>>>>

% -- Loads packages --
\usepackage[nott]{kpfonts}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[bf,singlelinecheck=off]{caption}
\usepackage[T1]{fontenc}
\usepackage{framed}
\usepackage{graphicx}
\usepackage[colorlinks=true]{hyperref} % web links
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{longtable}
\usepackage{multirow}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{placeins}
\usepackage{xcolor}

% Definition of colours for hyperref
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\definecolor{citecolour}{rgb}{0,0.8,0.2}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour,citecolor=citecolour}

\setcounter{secnumdepth}{4}

% -- New commands --
\newcommand{\relv}{release version 3.0}
\newcommand{\mol}{\textsc{OpenMolcas}}
\newcommand{\qcm}{\textsc{QCMaquis}}
\newcommand{\scine}{SCINE}
\newcommand{\hostp}{\mol}
\newcommand{\molbuild}{\texttt{my-Molcas-build}}
\newcommand{\molsrc}{\texttt{my-Molcas-src}}
\newcommand{\qcmsrc}{\texttt{my-QCMaquis-src}}
\newcommand{\qcmbuild}{\texttt{my-QCMaquis-build}}
\newcommand{\kwd}[1]{\texttt{#1}}

@@

@@

\lstset{
	numbers=left,
	numberstyle=\tiny,
	numberblanklines=true,
	basicstyle=\ttfamily\small,
	frame=single,
	showlines=true,
	flexiblecolumns=true,
	breaklines=true,
	postbreak={\llap{\textcolor{red}{$\lhook\joinrel\longrightarrow$\kern2.1em}}},
	escapechar=|
}

\title{A quick user guide to the\\ \scine-\qcm\ software suite for \mol}
\author{Alberto Baiardi, Robin Feldmann, Leon Freitag, Sebastian Keller, \\ Stefan Knecht, Yingjin Ma, Christopher Stein and Markus Reiher}
\publishers{ETH Z\"urich, Laboratorium f\"ur Physikalische Chemie, Vladimir-Prelog-Weg 2,\\ CH-8093 Z\"urich}
\date{\today}

@@

\vspace{2cm}

\noindent We kindly request that, for reproducibility reasons, any use of the \qcm\ software suite for density matrix renormalization group (DMRG) calculations in \mol\ that results in published material should cite the set-up steered by settings and warm-up procedures described in:

\begin{framed}
  \noindent {\textbf{Check for a preprint on \href{http://arxiv.org/}{arXiv.org} (to appear soon).}}\\
  \noindent Freitag~L.; Knecht~S.; Lindh~R.; Ma~Y.; Stein~C. J.; Reiher~M. \emph{in preparation}.
\end{framed}

\noindent The DMRG calculations are then conducted with the software \qcm\ that requires a citation.
It is described in the following paper:

\begin{framed}
\noindent Keller,~S.; Dolfi,~M.; Troyer,~M.; Reiher,~M. \emph{J. Chem. Phys.}
  \textbf{2015}, \emph{143}, 244118, \href{https://doi.org/10.1063/1.4939000}{doi:10.1063/1.4939000}.
\end{framed}

The reference works for the vibrational DMRG method are the following:

\begin{framed}
  \noindent Baiardi,~A.; Stein,~C.J.; Barone,~V.; Reiher,~M. \emph{J. Chem. Theory Comput.}
  \textbf{2017}, \emph{13}, 3764,   \href{https://doi.org/10.1021/acs.jctc.7b00329}{doi:10.1021/acs.jctc.7b00329}. \\
  \noindent Glaser,~N.; Baiardi~A.; Reiher,~M. ``Tensor Network States for Vibrational Spectroscopy'', \href{https://arxiv.org/abs/2109.08961}{arXiv.org}.
\end{framed}

The reference paper Nuclear-Electron All-Particle DMRG method (see Section~\ref{sec:preBO}) is the following:

\begin{framed}
  \noindent Muolo,~A.; Baiardi,~A.; Feldmann,~R.; Reiher,~M. \emph{J. Chem. Theory Phys.}
	\textbf{2020}, \emph{152}, 204103, \href{https://doi.org/10.1063/5.0007166}{doi:10.1063/5.0007166}.
\end{framed}

\qcm\ builds upon the ALPS MPS project. 
The \href{http://alps.comp-phys.org/static/mps\_doc/index.html}{ALPS MPS code} implements the DMRG algorithm for variational ground and low-lying excited state search as well as time evolution of arbitrary one- and two-dimensional models in a matrix-product-state representation.
They have been developed at ETH Zurich by Michele Dolfi and Bela Bauer in the group of Matthias Troyer with contributions from Sebastian Keller and Alexandr Kosenkov and at the University of Geneva by Timoth{\'e}e Ewart and Adrian Kantian in the group of Thierry Giamarchi.

@@

ALPS is a general open-source framework for the description of strongly correlated many-particle systems.

\begin{framed}
B. Bauer, \textit{et al.} (ALPS Collaboration), The ALPS project release 2.0: open source software for strongly correlated systems, \textit{J. Stat. Mech.} \textbf{2011} P05001. \href{http://dx.doi.org/10.1088/1742-5468/2011/05/P05001}{http://dx.doi.org/10.1088/1742-5468/2011/05/P05001}.
\end{framed}

@@

\pagenumbering{arabic}

\section{Introduction to the \qcm\ Software Suite}

\subsection{Overview and Goals}

The computational cost of traditional full CI and full CI-based multiconfigurational calculations (such as CASSCF) scales exponentially with the number of active orbitals.
Calculations beyond a certain limit (e. g. 18 electrons in 18 orbitals for CASSCF) become, therefore, prohibitively expensive. 
A possible approach to overcome the scaling limit is to employ methods such as density matrix renormalization group (DMRG) \cite{Schollwock_AnnPhys_densitymatrix_2011,Marti_ZPhysChem_Density_2010,Chan2011,Chan_JChemPhys_Matrix_2016,Baiardi2020_Review} to solve the full CI problem efficiently.

In DMRG, the wave function is represented in form of a matrix product state (MPS). 
The \qcm\ software suite solves the full CI problem approximately by means of an efficient optimization of an MPS wave function based on the so-called second-generation DMRG algorithm \cite{Keller_JChemPhys_efficient_2015}.
The quantum-chemical operators are represented as matrix product operators (MPOs), which provides the necessary flexibility to accommodate Abelian and non-Abelian symmetries, as well as the implementation of non-relativistic and relativistic quantum chemical Hamiltonians \cite{maquis-rel}, respectively, in a unified framework.
We leveraged the generality of the MPO/MPS-based framework to further extend \qcm\ to support Born--Oppenheimer\cite{Baiardi2017_vDMRG} and pre--Born--Oppenheimer\cite{Muolo2020_NEAP-DMRG} vibrational Hamiltonians.
We have implemented the special unitary group of degree 2 (SU(2)) in the MPO representation of the non-relativistic Hamiltonian to ensure
spin conservation \cite{Keller_JChemPhys_Spinadapted_2016}.
The current implementation of \qcm\ allows for efficient full-CI-type calculations of active space sizes well beyond capabilities (``CAS(18,18)") of standard CI approaches.

The \qcm\ software suite comes with a Fortran/Python-based interface \cite{interface} to the \mol\ \cite{Aquilante_JComputChem_Molcas_2016} framework, where we have implemented state-specific and state-average DMRG self-consistent field (DMRG-SCF) algorithms, the possibility to include solvent effects in DMRG calculations, analytic gradients for state-specific calculations, state interaction in a nonorthogonal basis,\cite{Knecht_JChemTheoryComput_nonorthogonal_2016} and n-electron valence state perturbation theory employing the DMRG reference wave function.\cite{Freitag_JChemTheoryComput_Multireference_2017}
These developments allow for, among others, structure optimizations, spin-orbit coupling calculations, evaluation of transition densities between non-orthogonal states for electronic and magnetic properties, as well as to account for dynamic correlation by means of the NEVPT2 approach.

\begin{framed}
  \large{\textbf{Advice:} The break-even point wrt computational cost for a DMRG-CI/-SCF calculation compared to a ``traditional" CAS-CI/-SCF calculation is approximately reached for a CAS(14,14) space. For active spaces smaller than CAS(14,14) we recommend to choose the conventional \mol\ CAS algorithm.}
\end{framed}

\subsection{What features are included in this release?}
\label{sec:features-v1.0}

@@

\subsection{Licenses}
\label{sec:license}

The \qcm{}-\mol{} interface and the NEVPT2 program are licensed under GNU LGPL, version 2.1.
The \qcm{} software suite and the ALPS library are distributed under \href{http://alps.comp-phys.org/static/software/applications/LICENSE.txt}{ALPS Application License version 1.0} and the \href{http://alps.comp-phys.org/static/software/ALPS/LICENSE.txt}{ALPS Library License version 1.0}.

\clearpage
\newpage
\section{Software Requirements \& Download}
\label{sec:install-req}

\subsection{Prerequisites}

In order to install either \qcm\ or the quantum chemistry package \mol\ with DMRG support through the \qcm\ software suite \cite{Keller_JChemPhys_efficient_2015,interface}, the following libraries/programs are required:

\begin{itemize}
 \item Git
 \item HDF5 (\url{http://www.hdfgroup.org/HDF5}) with Python2 bindings (\texttt{h5py})
 \item SZIP
 \item GNU Scientific Library GSL (\url{http://www.gnu.org/software/gsl/})
 \item CMake version $\geq$3.7 (\url{https://cmake.org/})
 \item Boost version $\geq$1.56 (\url{https://github.com/boostorg/boost})
 \item Python 3.x with \texttt{h5py} package installed\\
	   (optional, required only for MPSSI and distributed RDM calculations for NEVPT2 or the python tools)
\end{itemize}

\noindent Please make sure that these libraries/programs are available and their location is visible in your \texttt{\$PATH}\ and \texttt{\$LD\_LIBRARY\_PATH}, respectively.
Note that these libraries are \emph{NOT} part of the installation package of the \qcm\ software suite (see Section~\ref{sec:installation} for further details).

@@

If you want to download and install a standalone version of \qcm{}, it may be obtained from the github repository \url{https://github.com/qcscine/qcmaquis}.

\section{Installation}
\label{sec:installation}

\subsection{Installation of the \mol-\qcm{} interface}
\label{sec:install-qcm-mol}

The installation of \qcm\ has been tested for different operating systems and compiler/math libraries environments found in Section~\ref{subsec:supportedOS}.
While other combinations might work equally well they are \emph{not} officially supported.

\noindent The installation of the \qcm\ software suite within \mol\ will comprise several libraries which are \emph{automatically downloaded and installed} during the \mol\ build process provided that DMRG support within \mol\ is requested by the user.
The list of external libraries comprises:

\begin{itemize}
  \item Optionally: the NEVPT2 program
\end{itemize}

All of the above libraries will be installed locally in the user-defined build folder \texttt{my-Molcas-build} of \mol.

\subsubsection{Download and build folder setup}

Download \mol{} from the official \mol{} GitLab repository:

@@

\noindent Create a build folder \texttt{\molbuild} -- note that this folder \emph{does not} necessarily have to be a subfolder
of \texttt{my-Molcas-src} -- and change to this new folder:

\begin{verbatim}
  mkdir /path/to/my-Molcas-build && cd /path/to/my-Molcas-build
\end{verbatim}

\subsubsection{Configuration}
\label{subsubsec:configure}

@@

\paragraph{Configuration with the GNU compiler suite}
\label{sec:gnu-conf}$\;$\\

\noindent To configure \mol\ and \qcm\ with the GNU compiler suite type

\begin{verbatim}
  FC=gfortran CC=gcc CXX=g++ cmake -DDMRG=ON -DLINALG=MKL \
    /path/to/my-Molcas-src
\end{verbatim}

\noindent where we assumed (\texttt{-DLINALG=MKL}) that the MKL libraries are available (recommended option).
If the MKL libraries are not available internal math libraries of \mol\ can be requested with \texttt{-DLINALG=Internal} (default option for \texttt{LINALG} in \mol) or \texttt{-DLINALG=Runtime} (to use a system-wide BLAS or LAPACK installation). %Please also specify your name and e-mail address in the \texttt{-DQCMaquis\_NAME} and \texttt{-DQCMaquis\_EMAIL} CMake options, otherwise the download of \qcm{} will not work.

\begin{framed}
  \noindent\large{Note: We strongly recommend to configure \qcm\ (and \mol) with the Intel Math Kernel Library (MKL) to ensure the best numerical performance.
  See Section \ref{subsec:supportedOS}\ for further details.}
\end{framed}

\paragraph{Upgrading an existing \mol{} installation}
\label{sec:build-upgrade}$\;$\\

If you're upgrading from an \mol{} installation compiled with an older release of \qcm{}, you need to add an additional command line flag when invoking \texttt{cmake}, namely

\begin{verbatim}
  -DQCMaquis_UPDATE=ON
\end{verbatim}

This flag activates the \qcm{} update for the current and all the following \texttt{cmake} runs. To switch off \qcm{} updates, \texttt{cmake} may be run with

\begin{verbatim}
  -DQCMaquis_UPDATE=OFF
\end{verbatim}

@@

\paragraph{Additional options}$\;$\\

Following additional options can be switched on by passing additional flags to \texttt{cmake}:

\begin{itemize}
 \item \texttt{-DNEVPT2=ON}: Downloads and compiles the NEVPT2 program and its interface with \mol\ and \qcm.
 \item \texttt{-DOPENMP=ON}: Enables OpenMP parallelization for MOLCAS and NEVPT2 (see Section \ref{sec:parallel}).
       Note that \qcm{} is built with OpenMP support \emph{regardless} of this flag.
\end{itemize}

Other options to enable/disable various options in \mol{} may be passed to \texttt{cmake} at this step. Please consult the \mol{} manual for details.

% \paragraph{Additional \mol{} options}$\;$\\
% By running
% \begin{verbatim}
%  ccmake /path/to/my-Molcas-src
% \end{verbatim}
% in the build directory you may obtain a text user interface where several options for building \mol{}, along with the ones mentioned above, may be conveniently set. Use \texttt{ccmake} to set other various options for \mol{}: for details please consult the \mol{} manual.

\begin{framed}
  \noindent Note: it is currently discouraged to install the \qcm{}-\mol{} interface together with the \textsc{Block} or \textsc{CheMPS2} interfaces available in \mol{}.
  This setup has not been tested and may lead to an unusable installation.
  A fix allowing different interfaces to be installed at the same time is currently under development.
\end{framed}

@@

\noindent Installing an MPI-parallel version of \mol\ with DMRG support is possible although \qcm\ itself is by default shared-memory OMP but \emph{not} yet MPI-parallelized.
To configure \mol\ for an MPI-installation type

\begin{verbatim}
  FC=mpif90 CC=mpicc CXX=mpiCXX cmake -DDMRG=ON -DLINALG=MKL \
    -DQCMaquis_NAME="Your Name" -DQCMaquis_EMAIL="your@email.com" \
    /path/to/my-Molcas-src
\end{verbatim}

@@

A shared-memory OMP parallelized version of \mol\ can be activated with the option \texttt{-DOPENMP=ON}\ passed to \texttt{cmake} during the configuration step, for example

\begin{verbatim}
  FC=... CC=... CXX=... cmake -DDMRG=ON -DLINALG=MKL \
    -DQCMaquis_NAME="Your Name" -DQCMaquis_EMAIL="your@email.com" \
    /path/to/my-Molcas-src
\end{verbatim}

@@

After a successful configuration, type
\begin{verbatim}
  make
\end{verbatim}

\noindent or

\begin{verbatim}
  make -j8
\end{verbatim}
%
to compile \mol\ (in parallel on 8 cores) and install all its components in the build folder \texttt{\molbuild}.
In the same folder \qcm\ will be downloaded and installed.
The installation process thus requires a working internet connection.

@@

After having successfully passed the \qcm\ installation step as indicated by the CMake message

\begin{verbatim}
  [xxx%] Installation of QCMaquis, ALPS and Boost was successful!
\end{verbatim}

set up your \mol{} environment by setting \texttt{MOLCAS} environment variable

\begin{verbatim}
  export MOLCAS=/path/to/my-MOLCAS-build
\end{verbatim}

and then set up the \qcm{} runtime environment by running the following shell script

\begin{verbatim}
  source $MOLCAS/qcmaquis/bin/qcmaquis.sh
\end{verbatim}

You might want to include the above lines in your local \texttt{\$HOME/.bashrc} to set up your environment automatically every time your session is started.

@@

\qcm\ in \hostp\ is tested and officially supported for Linux (x86\_64) and MacOS X (10.12.x and 10.13.x ) operating systems.
We support GCC compilers starting from 5.x up to 9.x along with the MKL (Linux, x86\_64) and MKL/Accelerate (Mac OS X) math libraries, respectively, to successfully build the \qcm\ software suite within \hostp.
Note that other combinations might work equally well but they are not officially supported at present.

@@

To verify that your installation of \mol\ and \qcm\ was successful, run the following command

\begin{verbatim}
  pymolcas verify qcmaquis
\end{verbatim}

\noindent The message

\begin{verbatim}
  Verification has been completed
\end{verbatim}

\noindent indicates that all tests passed correctly and your installation is ready for production work.
\vspace{2ex}

@@

\begin{itemize}
  \item For issues regarding usage of \qcm\ in \mol{}, open an issue in the \mol{} bug tracker at \url{https://gitlab.com/Molcas/OpenMolcas/issues}
  \item For issues regarding the usage of \qcm{} standalone version, please open an issue on the \href{https://github.com/qcscine/qcmaquis}{\qcm\ GitHub repository}.
\end{itemize}

Any issue will be dealt with by one of the authors, although no guarantee on the promptness of response is given.
In general, serious bugs that have been reported and fixed will lead to a new patch of the \qcm\ program, announced and distributed via the \href{https://github.com/qcscine/qcmaquis}{\qcm\ GihHub repository}.
